Random Matrix Theory Tools for the Analysis of Functional Magnetic Resonance Imaging Examinations

### 2021-09-13 paper/latex/rmt.tex
formatting

\begin{center}
\texttt{\textbf{INSERT TABLE 1 HERE}}
\end{center}

@@

\subsubsection{Dataset 1}
Dataset 1 (referred to in Fig. and tables with the label "PSYCH") has detailed and extensive
information available in \cite{gorgolewskiHighResolution7Tesla2015}. Subjects completed a battery of
psychological measurements, and various resting-state fMRI (rs-fMRI) images were obtained. Only the
whole-brain 7-Tesla rs-fMRI scans were used for the current paper, to make the results comparable to
the other included datasets. For the purpose of keeping with the "attention" theme of the analyses
in this paper, three pairings of subgroups were constructed via z-score splits of the various
sum-scores of scale items involving concepts closely related to attention.

\paragraph{Trait Attention}
The PANAS-X (\cite{watsonPANASXManualPositive1994}) is a 9-point Likert-type self-report measure
that queries the extent to which subjects agree particular affective terms have applied to them over
a recent time period (e.g. a few weeks). An ad hoc (i.e. face-valid) sum-score was constructed from
the scale items: "attentive", "alert", "interested", "concentrating", and reverse-scored items
"sluggish", "tired", "sleepy", "drowsy". Subjects with a z-score greater / less than zero on this
sum-score were placed in the high / low "trait attention" category

\paragraph{Task Attention}
The Conjunctive Continuous Performance Task (CCPT;
\cite{shalevConjunctiveContinuousPerformance2011}) is a behavioural measure of sustained attention
where subjects must react only if they see a red square among other different coloured figures. For
this paper, a sum score was constructed by subtracting the number of false positive and false
negative responses from the number of success responses, and subjects with a z-score greater / less
than zero on this scale were placed in high / low "task attention" category.

\paragraph{Vigilance}
The Mini New York cognition questionnaire assesses various elements of the subject's cognitive
state, e.g. "my thoughts involved other people", "I thought about something positive" (translated
from German) \cite{gorgolewskiCorrespondenceIndividualDifferences2014}. One question assesses
alertness ("I was fully awake" / "war ich vollkommen wach") and was used to create high / low
"vigilance" conditions in the same manner as the previous two splits.

\subsubsection{Dataset 2}
Dataset 2 (referred to in Fig. and tables with the label "OSTEO") is detailed in
\cite{tetreaultBrainConnectivityPredicts2016}. Subjects included healthy controls ("nopain"
condition) and individuals with knee osteoarthritis ("pain" condition) treated for 2 weeks with
either placebo ("pain" condition) or duloxetine ("duloxetine" condition). An additional grouping was
also created with both the placebo and duloxetine data ("allpain" condition), allowing for six
subgroup comparisons. All subjects were imaged with whole-brain rs-fMRI. Scan parameters are
documented in Table I. The use of rs-fMRI and RMT for examining pain makes this dataset
complementary to \cite{matharooSpontaneousBackpainAlters2020}.

\subsubsection{Dataset 3}
Dataset 3  (referred to here with the label "REFLECT", due to the "reflective” nature of the
cognitive tasks) is described in full detail by \cite{dupreMultiechoFMRIReplication2016}. Both
multi-echo rs- and task-fMRI scans were conducted with scan parameters summarized in Table II. Tasks
involved viewing various affective images (see \cite{sprengPatternsBrainActivity2010} for details)
and performing tasks related to autobiographical memory, prospection, or theory-of-mind reasoning.
Continuing with the theme of broadly comparing high versus low attention throughout this paper, the
different tasks were simply treated generically as "attending", in contrast to the non-attentive
resting-state scans. This resulted in four task state scans and one resting-state scan per subject.

In order to combine the multi-echo data into a single, usable matrix, two methods were employed. In
the "summed" condition, echoes were combined via summation, a method which has value in providing an
image with overall increased contrast (\cite{kunduMultiechoFMRIReview2017}). In the "interleave"
condition, echoes were interleaved at each voxel: that is, if we denote the voxel's ith signal value
from echo j as , then the interleaved voxel of the new image has the time-series \([t_{1,1},
t_{2,1}, t_{3,1}, \dots, t_{1,i}, t_{2,i}, t_{3,i}, \dots, t_{1,n}, t_{2,n}, t_{3,n}]\).

@@

\subsubsection{Dataset 4}
Dataset 4 (referred to in Fig. and tables with the label "PARK" or "PARKINSONS") is documented and
described in detail in \cite{madhyasthaDynamicConnectivityRest2015}. Scan parameters are documented
in Table I. Subjects with non-demented Parkinson’s Disease (PD) and healthy controls performed a
number of repetitions of the Attention Network Task (ANT)
(\cite{fanActivationAttentionalNetworks2005}). Subgroup details are available in Table I. As the ANT
task demands high attention, it is not as straightforward to create "high" or "low" attention
groups. Instead, since attention presumably is high and varies little in this dataset, and subgroups
instead differ by virtue of the presence or absence of PD, any results here serve as a useful
comparison point for the results from studies that should clearly involve attentional differences.
This dataset also included extensively preprocessed scans, which were included in analyses to
examine the effects of preprocessing on the extracted RMT features.

\subsubsection{Dataset 5}
Dataset 5 (referred to in Fig. and tables with the label "LEARN") is documented and described in
detail in \cite{schapiroHumanHippocampalReplay2018}. In this study, both rs-fMRI and task fMRI scans
were available for all subjects, allowing for within-subject comparisons. The complex task involved
multiple phases, wherein subjects were exposed to named, black-and-white outlined satellite images
related by family-resemblance of the components (see Fig. 1 of
\cite{schapiroHumanHippocampalReplay2018}). After this learning task, subjects were tested on their
memory of the various learned features. Scans corresponding to both the initial learning and
subsequent testing phase were included in the high-attention "task" subgroup for the purposes of the
analysis of this paper. Resting state scans were included in the low-attention "rest" subgroup.

@@

A number of statistics (or "spectral observables"; \cite{mehtaRandomMatrices2004}) can then be
computed from this unfolded spectrum. In the literature, the two most common points of comparison
for these unfolded features are typically the features of the Poisson or Gaussian Diagonal Ensemble
(GDE) and the GOE (see e.g.
\cite{santhanamStatisticsAtmosphericCorrelations2001,jalanUncoveringRandomnessSuccess2014,bandyopadhyayUniversalityComplexNetworks2007,agrawalQuantifyingRandomnessProtein2014,raiRandomnessPreservedPatterns2015,
sebaRandomMatrixAnalysis2003,wangSpectralPropertiesTemporal2015,wangRandomMatrixTheory2016}). The
GOE (or GDE) consists of real, symmetric matrices where entries (or only diagonal entries) are
normal random variables with zero mean and unit variance. When the empirically observed features are
close to those of the GOE, this is usually interpreted as indicating a high degree of randomness and
statistical independence in the data, and when the observed features are closer to those expected in
the GDE or Poisson case, this is interpreted as indicating a high degree of coherence or dependence
(\cite{santhanamStatisticsAtmosphericCorrelations2001,jalanUncoveringRandomnessSuccess2014,bandyopadhyayUniversalityComplexNetworks2007,agrawalQuantifyingRandomnessProtein2014,raiRandomnessPreservedPatterns2015,
sebaRandomMatrixAnalysis2003,wangSpectralPropertiesTemporal2015,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}).

@@

In addition to the unfolded features suggested by RMT, there are some potential "raw" features that
do not require an unfolding procedure. The entire eigenvalue spectrum likely contains important
information related to the system, and is also investigated as an RMT-inspired feature.

@@

The sorted, raw eigenvalues, however, more consistently contained information that allowed some
separation of subgroups. Averaging across all datasets and comparisons, the raw eigenvalues
consistently demonstrated predictive utility (Tables S7--8). As visible in Fig. S13, the eigenvalues
had the most consistent predictive value for distinguishing individuals with Parkinson's from
controls, or distinguishing those with osteopathic pain taking duloxetine from controls ("nopain" in
Fig. S13). The raw eigenvalues sometimes had predictive value for the PSYCH data, most notably in
comparing those differing in self-reported vigilance and high vs. low scorers in the attention task
(Table S7--8).

@@

\paragraph{Role of Classifer}
The raw eigenvalues generally had the most predictive utility when employing either an SVM
classifier or KNN classifier with k > 3 (Table S11, first row). When the raw eigenvalues were first
min-max normalized prior to prediction, an SVM classifier was clearly the most consistent in
predicting subgroup memberships (Table S12), however without normalization, KNN with k = 8 or k = 9
generally best predicted subgroup membership across datasets (Table S11).

@@

\section{Discussion}
In this study, certain features inspired by RMT and derived from the sorted eigenvalues of the full,
whole-brain voxelwise correlation matrix were found to contain information with the potential to
meaningfully differentiate between different neuropsychological phenomena. In addition to evidence
from previous studies [26], [33], [34], this suggests RMT may be a useful analytic and theoretical
tool for understanding functional connectivity.

@@

For comparison, for small datasets of rs- or task-fMRI data, binary classification using whole-brain
features is challenging. For subject level classification, state of the art deep learning methods
improve upon guessing by 17\% for autism [71] or 3-30\% for ADHD [72], 16\% for severe depression [73],
and 23\% for obsessive compulsive disorder [74]. Manual feature engineering with more separable
conditions (e.g. schizophrenia) can result in classification accuracies well above 90\% [75], and
with very large data, sophisticated custom feature extraction methods can achieve near perfect
accuracies at classifying task vs. rest [76]. However, for functional connectivity data and with
classical machine learning algorithms (such as SVM) we in general only expect large prediction
accuracies (e.g. greater than 80\%) when the group functional connectivities are already strongly
separated (e.g. Cohen’s d > 1.0 [77]).

@@

\subsection{Confounds}
While some of the RMT features allowed some classifiers to nearly perfectly separate conditions for
the REFLECT dataset, this should almost certainly be regarded as an artifact of the differing scan
parameters between task and rest conditions (Table II). The size of the largest eigenvalue of a
correlation matrix is primarily determined by such differences in dimensions [78], as are the noise
ratios, which include these dimensions in their definitions [29], [30]. Likewise, similar strong
separation was observed in the PARK data when comparing the minimally pre-processed to fully
pre-processed data (Fig. S1-4). The pre-processing in this case involved numerous steps (spike
removal, spatial smoothing, co-registration to larger T1 image [45]) which will also almost
certainly alter the size of the largest eigenvalue.

@@

However, that there are not consistently strong differences in these unidimensional features between
subgroups across various datasets (Fig. S8-10) also suggests that employing these features in, e.g.
noise-reduction procedures (such as [29], [30]), is unlikely to introduce bias into subsequent
analyses. It perhaps also suggests these scalar features might be useful for quality control within
a dataset: if there are differences between subgroups or individuals on one of the individual RMT
features, then perhaps this is cause for more careful investigation.

@@

We generally find different results from previous studies [26], [33], [34] which have used the
spectral rigidity or level number variance to differentiate between attention conditions. As [26]
extracted rigidity curves from the correlations of electroencephalographic (EEG) signals, and [34]
and [33] extracted spectral observables from the correlations of the mean signals of anatomical
ROIs, these differences are to be expected. A single voxel signal amalgamates information from far
fewer neurons than an EEG or mean ROI signal, and it is possible that any broad connections between
RMT spectral observables and attention arise only at higher levels of organization, or after certain
pre-processing steps (e.g. spatial smoothing) not performed in this study.

>>>>

\begin{table}[h!]
\small
\centering
\begin{tabular}{ c c c c c c c }
\hline
\textbf{ID}  & \textbf{FOV (mm)}  & \textbf{Dimensions}  & \textbf{Voxel size (mm)} & \textbf{TR} & \textbf{Volumes} & \textbf{Time (s)} \\
\hline
PSYCH   &  192              & 128 x 128 x 70 & 1.5 x 1.5 x 1.5 & 3.0 & 300 & 15 \\
OSTEO   &  -                & 64 x 64 x 36   & 3.4 x 3.4 x 3.0 & 2.5 & 300 & 12.5 \\
REFLECT &  210              & 72 x 72 x 46   & 3.0 x 3.0 x 3.0 & 2.5 & 204 & 10.2 \\
REFLECT &  240              & 64 x 64 x 33   & 3.8 x 3.8 x 3.8 & 2.0 & 260 & 8.7 \\
PARK    &  240 x 240 x 129  & 80 x 80 x 43   & 3.0 x 3.0 x 3.0 & 2.4 & 149 & 5.96 \\
LEARN   &  -                & 64 x 64 x 36   & 3.0 x 3.0 x 3.0 & 2.0 & 196 & 6.5 \\
\hline
\end{tabular}
\caption{
    ID = Identifier for paper. FOV = Field of View. TR = Time of Repetition (seconds).
    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
    indicate P slices with dimensions M x N.
}
\label{table:1}
\end{table}

@@

\begin{table}[h!]
\small
\centering
\begin{tabular}{ l l c c c }
\hline
\textbf{ID}  & \textbf{Subgroup / Condition}  & \textbf{Subjects}  & \textbf{Task} & \textbf{Scans per Subject} \\
\hline
PSYCH   &  high\_vigilance    & 12 / 11 & resting-state                    & 2 \\
        &  low\_vigilance     & 10 / 11 & resting-state                    & 2 \\
        &  trait\_attend      & 12 / 15 & resting-state                    & 2 \\
        &  trait\_non-attend  & 10 / 7  & resting-state                    & 2 \\
        &  task\_attend       & 13 / 15 & resting-state                    & 2 \\
        &  task\_nonattend    &  9 / 7  & resting-state                    & 2 \\
\hline
LEARN   &  task              & 24      & learn image sketches & 16 \\
        &  rest              & 24      & resting-state                    & 2 \\
\hline
REFLECT &  control (no pain) & 20      & resting-state                    & 1 \\
        &  duloxetine        & 19      & resting-state                    & 1 \\
        &  placebo           & 37      & resting-state                    & 1 \\
        &  allpain (duloxetine + placebo) & 56         & resting-state    & 1 \\
\hline
PARK    &  Parkinson's       & 17      & ANT                              & 6 \\
        &  control           & 17      & ANT                              & 6 \\
\hline
ECHO    &  summed/interleave & 31      & resting-state                    & 1 \\
        &  summed/interleave & 31      & imagery/reflection               & 4 \\
\hline
\end{tabular}
\caption{
    ID = Identifier for paper. FOV = Field of View. TR = Time of Repetition (seconds).
    Time = total duration (minutes) of each scan. Dimensions listed as M x N x P,
    indicate P slices with dimensions M x N.
}
\label{table:2}
\end{table}

\subsubsection{Dataset 1}
Dataset 1 (referred to in Fig. and tables with the label ``PSYCH'') has detailed and extensive
information available in \cite{gorgolewskiHighResolution7Tesla2015}. Subjects completed a battery of
psychological measurements, and various resting-state fMRI (rs-fMRI) images were obtained. Only the
whole-brain 7-Tesla rs-fMRI scans were used for the current paper, to make the results comparable to
the other included datasets. For the purpose of keeping with the ``attention'' theme of the analyses
in this paper, three pairings of subgroups were constructed via z-score splits of the various
sum-scores of scale items involving concepts closely related to attention.

\paragraph{Trait Attention}
The PANAS-X (\cite{watsonPANASXManualPositive1994}) is a 9-point Likert-type self-report measure
that queries the extent to which subjects agree particular affective terms have applied to them over
a recent time period (e.g. a few weeks). An ad hoc (i.e. face-valid) sum-score was constructed from
the scale items: ``attentive'', ``alert'', ``interested'', ``concentrating'', and reverse-scored items
``sluggish'', ``tired'', ``sleepy'', ``drowsy''. Subjects with a z-score greater / less than zero on this
sum-score were placed in the high / low ``trait attention'' category

\paragraph{Task Attention}
The Conjunctive Continuous Performance Task (CCPT;
\cite{shalevConjunctiveContinuousPerformance2011}) is a behavioural measure of sustained attention
where subjects must react only if they see a red square among other different coloured figures. For
this paper, a sum score was constructed by subtracting the number of false positive and false
negative responses from the number of success responses, and subjects with a z-score greater / less
than zero on this scale were placed in high / low ``task attention'' category.

\paragraph{Vigilance}
The Mini New York cognition questionnaire assesses various elements of the subject's cognitive
state, e.g. ``my thoughts involved other people'', ``I thought about something positive'' (translated
from German; \cite{gorgolewskiCorrespondenceIndividualDifferences2014}). One question assesses
alertness (``I was fully awake'' / ``war ich vollkommen wach'') and was used to create high / low
``vigilance'' conditions in the same manner as the previous two splits.

\subsubsection{Dataset 2}
Dataset 2 (referred to in Fig. and tables with the label ``OSTEO'') is detailed in
\cite{tetreaultBrainConnectivityPredicts2016}. Subjects included healthy controls (``nopain''
condition) and individuals with knee osteoarthritis (``pain'' condition) treated for 2 weeks with
either placebo (``pain'' condition) or duloxetine (``duloxetine'' condition). An additional grouping was
also created with both the placebo and duloxetine data (``allpain'' condition), allowing for six
subgroup comparisons. All subjects were imaged with whole-brain rs-fMRI. Scan parameters are
documented in Table I. The use of rs-fMRI and RMT for examining pain makes this dataset
complementary to \cite{matharooSpontaneousBackpainAlters2020}.

\subsubsection{Dataset 3}
Dataset 3  (referred to here with the label ``REFLECT'', due to the ``reflective” nature of the
cognitive tasks) is described in full detail by \cite{dupreMultiechoFMRIReplication2016}. Both
multi-echo rs- and task-fMRI scans were conducted with scan parameters summarized in Table II. Tasks
involved viewing various affective images (see \cite{sprengPatternsBrainActivity2010} for details)
and performing tasks related to autobiographical memory, prospection, or theory-of-mind reasoning.
Continuing with the theme of broadly comparing high versus low attention throughout this paper, the
different tasks were simply treated generically as ``attending'', in contrast to the non-attentive
resting-state scans. This resulted in four task state scans and one resting-state scan per subject.

In order to combine the multi-echo data into a single, usable matrix, two methods were employed. In
the ``summed'' condition, echoes were combined via summation, a method which has value in providing an
image with overall increased contrast (\cite{kunduMultiechoFMRIReview2017}). In the ``interleave''
condition, echoes were interleaved at each voxel: that is, if we denote the voxel's ith signal value
from echo j as , then the interleaved voxel of the new image has the time-series \([t_{1,1},
t_{2,1}, t_{3,1}, \dots, t_{1,i}, t_{2,i}, t_{3,i}, \dots, t_{1,n}, t_{2,n}, t_{3,n}]\).

@@

\subsubsection{Dataset 4}
Dataset 4 (referred to in Fig. and tables with the label ``PARK'' or ``PARKINSONS'') is documented and
described in detail in \cite{madhyasthaDynamicConnectivityRest2015}. Scan parameters are documented
in Table I. Subjects with non-demented Parkinson’s Disease (PD) and healthy controls performed a
number of repetitions of the Attention Network Task (ANT)
(\cite{fanActivationAttentionalNetworks2005}). Subgroup details are available in Table I. As the ANT
task demands high attention, it is not as straightforward to create ``high'' or ``low'' attention
groups. Instead, since attention presumably is high and varies little in this dataset, and subgroups
instead differ by virtue of the presence or absence of PD, any results here serve as a useful
comparison point for the results from studies that should clearly involve attentional differences.
This dataset also included extensively preprocessed scans, which were included in analyses to
examine the effects of preprocessing on the extracted RMT features.

\subsubsection{Dataset 5}
Dataset 5 (referred to in Fig. and tables with the label ``LEARN'') is documented and described in
detail in \cite{schapiroHumanHippocampalReplay2018}. In this study, both rs-fMRI and task fMRI scans
were available for all subjects, allowing for within-subject comparisons. The complex task involved
multiple phases, wherein subjects were exposed to named, black-and-white outlined satellite images
related by family-resemblance of the components (see Fig. 1 of
\cite{schapiroHumanHippocampalReplay2018}). After this learning task, subjects were tested on their
memory of the various learned features. Scans corresponding to both the initial learning and
subsequent testing phase were included in the high-attention ``task'' subgroup for the purposes of the
analysis of this paper. Resting state scans were included in the low-attention ``rest'' subgroup.

@@

A number of statistics (or ``spectral observables''; \cite{mehtaRandomMatrices2004}) can then be
computed from this unfolded spectrum. In the literature, the two most common points of comparison
for these unfolded features are typically the features of the Poisson or Gaussian Diagonal Ensemble
(GDE) and the GOE (see e.g.
\cite{santhanamStatisticsAtmosphericCorrelations2001,jalanUncoveringRandomnessSuccess2014,bandyopadhyayUniversalityComplexNetworks2007,agrawalQuantifyingRandomnessProtein2014,raiRandomnessPreservedPatterns2015,
sebaRandomMatrixAnalysis2003,wangSpectralPropertiesTemporal2015,wangRandomMatrixTheory2016}). The
GOE (or GDE) consists of real, symmetric matrices where entries (or only diagonal entries) are
normal random variables with zero mean and unit variance. When the empirically observed features are
close to those of the GOE, this is usually interpreted as indicating a high degree of randomness and
statistical independence in the data, and when the observed features are closer to those expected in
the GDE or Poisson case, this is interpreted as indicating a high degree of coherence or dependence
(\cite{santhanamStatisticsAtmosphericCorrelations2001,jalanUncoveringRandomnessSuccess2014,bandyopadhyayUniversalityComplexNetworks2007,agrawalQuantifyingRandomnessProtein2014,raiRandomnessPreservedPatterns2015,
sebaRandomMatrixAnalysis2003,wangSpectralPropertiesTemporal2015,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}).

@@

In addition to the unfolded features suggested by RMT, there are some potential ``raw'' features that
do not require an unfolding procedure. The entire eigenvalue spectrum likely contains important
information related to the system, and is also investigated as an RMT-inspired feature.

@@

The sorted, raw eigenvalues, however, more consistently contained information that allowed some
separation of subgroups. Averaging across all datasets and comparisons, the raw eigenvalues
consistently demonstrated predictive utility (Tables S7--8). As visible in Fig. S13, the eigenvalues
had the most consistent predictive value for distinguishing individuals with Parkinson's from
controls, or distinguishing those with osteopathic pain taking duloxetine from controls (``nopain'' in
Fig. S13). The raw eigenvalues sometimes had predictive value for the PSYCH data, most notably in
comparing those differing in self-reported vigilance and high vs. low scorers in the attention task
(Table S7--8).

@@

\paragraph{Role of Classifer}
The raw eigenvalues generally had the most predictive utility when employing either an SVM
classifier or KNN classifier with \(k > 3\) (Table S11, first row). When the raw eigenvalues were
first min-max normalized prior to prediction, an SVM classifier was clearly the most consistent in
predicting subgroup memberships (Table S12), however without normalization, KNN with \(k = 8\) or
\(k = 9\) generally best predicted subgroup membership across datasets (Table S11).

@@

\section{Discussion}
In this study, certain features inspired by RMT and derived from the sorted eigenvalues of the full,
whole-brain voxelwise correlation matrix were found to contain information with the potential to
meaningfully differentiate between different neuropsychological phenomena. In addition to evidence
from previous studies
(\cite{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020})
this suggests RMT may be a useful analytic and theoretical tool for understanding functional
connectivity.

@@

For comparison, for small datasets of rs- or task-fMRI data, binary classification using whole-brain
features is challenging. For subject level classification, state of the art deep learning methods
improve upon guessing by 17\% for autism (\cite{bengs4DSpatioTemporalDeep2020}) or 3-30\% for ADHD
(\cite{riazDeepFMRIEndtoendDeep2020}), 16\% for severe depression
(\cite{ramasubbuAccuracyAutomatedClassification2016}), and 23\% for obsessive compulsive disorder
(\cite{takagiNeuralMarkerObsessiveCompulsive2017}). Manual feature engineering with more separable
conditions (e.g. schizophrenia) can result in classification accuracies well above 90\%
(\cite{duHighClassificationAccuracy2012}), and with very large data, sophisticated custom feature
extraction methods can achieve near perfect accuracies at classifying task vs. rest
(\cite{zhangCharacterizingDifferentiatingTaskbased2016}). However, for functional connectivity data
and with classical machine learning algorithms (such as SVM) we in general only expect large
prediction accuracies (e.g. greater than 80\%) when the group functional connectivities are already
strongly separated (e.g. Cohen’s d > 1.0; \cite{dansereauStatisticalPowerPrediction2017}).

@@

\subsection{Confounds}
While some of the RMT features allowed some classifiers to nearly perfectly separate conditions for
the REFLECT dataset, this should almost certainly be regarded as an artifact of the differing scan
parameters between task and rest conditions (Table II). The size of the largest eigenvalue of a
correlation matrix is primarily determined by such differences in dimensions
(\cite{yinLimitLargestEigenvalue1988}), as are the noise ratios, which include these dimensions in
their definitions (\cite{veraartDiffusionMRINoise2016,veraartDenoisingDiffusionMRI2016}). Likewise,
similar strong separation was observed in the PARK data when comparing the minimally pre-processed
to fully pre-processed data (Fig. S1-4). The pre-processing in this case involved numerous steps
(spike removal, spatial smoothing, co-registration to larger T1 image;
\cite{madhyasthaDynamicConnectivityRest2015}) which will also almost certainly alter the size of the
largest eigenvalue.

@@

However, that there are not consistently strong differences in these unidimensional features between
subgroups across various datasets (Fig. S8-10) also suggests that employing these features in, e.g.
noise-reduction procedures (such as
\cite{veraartDiffusionMRINoise2016,veraartDenoisingDiffusionMRI2016}), is unlikely to introduce bias
into subsequent analyses. It perhaps also suggests these scalar features might be useful for quality
control within a dataset: if there are differences between subgroups or individuals on one of the
individual RMT features, then perhaps this is cause for more careful investigation.

@@

We generally find different results from previous studies
(\cite{sebaRandomMatrixAnalysis2003,wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}
which have used the spectral rigidity or level number variance to differentiate between attention
conditions. As \cite{sebaRandomMatrixAnalysis2003} extracted rigidity curves from the correlations
of electroencephalographic (EEG) signals, and \cite{wangRandomMatrixTheory2016} and
\cite{matharooSpontaneousBackpainAlters2020} extracted spectral observables from the correlations of
the mean signals of anatomical ROIs, these differences are to be expected. A single voxel signal
amalgamates information from far fewer neurons than an EEG or mean ROI signal, and it is possible
that any broad connections between RMT spectral observables and attention arise only at higher
levels of organization, or after certain pre-processing steps (e.g. spatial smoothing) not performed
in this study.
